---
title: "Busselton Lipid"
author: "nminaee"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r,echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(readr)
library(reshape2)
library(ggplot2)
library(ggrepel)
library(tibble)
library(ggpubr)
library(mva.plots)
library(ca)
library(tidyr)
library(tidyverse)
library(corrplot)
library(GGally)
library(stats)
library(pheatmap)
library(cluster)
library(gridExtra)
library(ggpubr)
library(htmlTable)
library(broom)
```


```{r, echo=FALSE}
# Create a function to set square aspect ratio
set_square_aspect_ratio <- function(plot) {
  plot + theme(aspect.ratio = 1)
}


continuousPalette<- c(
                        "#0000CC",
                        "#0000FF",
                        "#0055FF",
                        "#00AAFF",
                        "#00FFFF",
                        "#2BFFD5",
                        "#55FFAA",
                        "#80FF80",
                        "#AAFF55",
                        "#D4FF2B",
                        "#FFFF00",
                        "#FFAA00",
                        "#FF5500",
                        "#FF0000",
                        "#CC0000")

txt2S <- function(col) {
  newCol <- lapply(col, function(i) {
    if (!is.na(i)){
      i <- abs(as.numeric(i))
      if (i < 0.05 & i >= 0.01) {"*"
      } else if (i < 0.01 & i >= 0.001) {
        "**"
      } else if (i < 0.001) {
        "***"
      } else {
        ""
      }
    } else {
      NA
    }
  })
  return(unlist(newCol))
}


# Function to perform univariate t-test
univariate_test <- function(data) {
  t.test(value ~ group_var, data = data) %>% 
    broom::tidy()
}



perform_stat_tests<-function(tdf_tbl,group = NULL,var, method, tbl1){
  # Handle missing or infinite values
  data_pval <- tdf_tbl %>%
    mutate(across(-{{group}}, ~replace(., is.nan(.) | is.infinite(.), 0)))%>%
    rename(group = {{group}})%>%
    dplyr::select(group,var)
  
  
  # Switch between ANOVA and Wilcoxon tests based on the method argument
  p_model <- switch(method,
                    "anova" = lapply(data_pval %>% select(-group), function(x) {
                      aov(x ~ data_pval$group)
                    }),
                    "wilcox" = {
                      if (length(unique(data_pval$group)) > 2) {
                        stop("Wilcoxon test does not support more than 2 groups naturally.")
                      } else {
                        lapply(data_pval %>% select(-group), function(x) {
                          wilcox.test(x ~ data_pval$group)
                        })
                      }
                    },
                    stop("Unsupported method type provided")
  )
  
  # Generate summaries of the model results
  model_summaries <- lapply(p_model, summary)
  
  
  # Extracting results and performing post-hoc tests where applicable
  summary_list <- list()
  
  for(i in names(model_summaries)){
    variable = data.frame(Variable =i)
    variable$Pvalue = txt2S(model_summaries[[i]][[1]]['Pr(>F)'][1,])
    tukey_p<-as.data.frame(TukeyHSD(p_model[[i]])[[1]])
    tukey_p$comparison<-rownames(tukey_p)
    tukey_p$`p adj`<-txt2S(tukey_p$`p adj`)
    tukey_p<-data.frame(t(tukey_p[,c("comparison","p adj")]))
    colnames(tukey_p)<-tukey_p[1,]
    tukey_p<-tukey_p[-1,]
    variable<-cbind(variable,tukey_p)
    summary_list[[i]]<-variable
    rm(variable,tukey_p)
  }
  
  # Combine all individual summaries into a single data frame
  summary_list <- do.call(rbind, summary_list)
  
  # Merge the summarized data frame with the existing table `tbl1`
  tbl1 <- merge(tbl1, summary_list, by = "Variable")
  
  return(tbl1)
}
```


# Load the data

MS Lipid 
```{r, echo=FALSE, message= FALSE, warning=FALSE}
load("/Users/novia/Desktop/Diabetes_Obesity/Busselton/LipidMS/Busselton_PLA_MS_Lipids.daE")

lipidData <- Busselton_PLA_MS_Lipids@.Data
lipidMeta <- Busselton_PLA_MS_Lipids@obsDescr[[1]] # only need this part to retrieve sampleID that has been matched to lipidData; do not use this metadata - use the Busselton metadata
# only use lipid data, use the Busselton metadata for the metadata. 

l <- cbind(lipidData, lipidMeta)

l <- l %>% select(1:895, sampleID)

#rm(Busselton_PLA_MS_Lipids, lipidData, lipidMeta)
```

Lipoprotein
```{r, echo=FALSE, message= FALSE, warning=FALSE}
b <- read.csv("/Users/novia/Desktop/Diabetes_Obesity/Busselton/Matched_Bus_anno_lipo_allcolumns_NM_13052024.csv", header = TRUE)

#load("Busselton_Final_forPaper2_23May2024.rda") #buss4 - final lipo data + meta that's used for Paper 2 draft. buss4 is the simplified version of b.  
# NOTE: b and buss4 sampleID matched (08/07/2024). 

#table(b$Summary_Type2_Diabetes_idx, b$Diabetes_)
# check if there's any discrepancies between DM and diabetes group.
### NOTE: Diabetes_ and Summary_Type2_Diabetes give roughly the same number of overall NO DM vs DM. 
### BUT Diabetes_ = 1 includes uncertain DM1 diagnosis. So here onwards it's safer to use Summary_Type2_Diabetes_idx. 
### DM column in buss4 reflects Summary_Type2_Diabetes column in b. 
```


```{r, echo=FALSE, message= FALSE, warning=FALSE}
# selecting only timepoint 1 (T1) samples from lipid data to match lipo data and meta. 
idx<-which(l$sampleID %in% b$sampleID)
lipid <- l[which(l$sampleID %in% b$sampleID),]
lipid <- lipid[match(b$sampleID, lipid$sampleID),]

# checking if matching works
#table(lipid$sampleID == b$sampleID) # yes 1976 matched. 
# NA checking
na_rows <- lipid[!complete.cases(lipid), ] #41 NAs

lipid <- lipid %>% filter(!lipid$sampleID %in% na_rows$sampleID) #lipid data
#1935

lipo <- b %>% filter(!b$sampleID %in% na_rows$sampleID) #lipo data

#table(lipid$sampleID == lipo$sampleID) #1935 Phew!

lipo_names<-colnames(lipo)[1:112]
lipid_names<-colnames(lipid)[1:895]
lipid_class<-unique(sub("\\(.*$", "", lipid_names))

combined <- merge(lipid, lipo, by = "sampleID")

combined$taking_lipid_med <- ifelse(grepl("lipid|HMG-CoA|statin", combined$an_med, ignore.case = TRUE), 1, 0)
combined$taking_any_med <- ifelse(is.na(combined$an_med), 0, 1)

combined <- combined %>% relocate(sampleID,.before = "TPTG")

#save(combined,file="Busselton_LipidMS_matchedwLipo_plusmeds_15Jul2024.rda")

#rm(idx, na_rows, b, l, lipid, lipo)
```


```{r, echo = FALSE, message=FALSE, warning=FALSE}
#load("Busselton_LipidMS_matchedwLipo_plusmeds_15Jul2024.rda")
```


# Split 1: Healthy = not diabetic (DM status = No DM) but can include other conditions.
```{r, echo=FALSE, message= FALSE, warning=FALSE}
healthy_nodm <- combined %>% filter(combined$DM == "NO DM") #1784

healthy_nodm_hw <- healthy_nodm %>% filter(healthy_nodm$BMI.weight == "Healthy") #491
healthy_nodm_ob <- healthy_nodm %>% filter(healthy_nodm$BMI.weight == "Obese") #472

healthy_nodm_NOmeds_hw <- healthy_nodm_hw %>% filter(healthy_nodm_hw$taking_lipid_med == 0) #461
healthy_nodm_meds_hw <- healthy_nodm_hw %>% filter(healthy_nodm_hw$taking_lipid_med == 1) #30

healthy_nodm_NOmeds_ob <- healthy_nodm_ob %>% filter(healthy_nodm_ob$taking_lipid_med == 0) #385
healthy_nodm_meds_ob <- healthy_nodm_ob %>% filter(healthy_nodm_ob$taking_lipid_med == 1) #87

```


## Heatmap with Clustering
### Overall BMI - Healthy-weight vs Obese
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load datasets
data1 <- healthy_nodm_hw %>% select(1:895)  
data2 <- healthy_nodm_ob %>% select(1:895) 

# Calculate the correlation matrices
cor1 <- cor(data1)
cor2 <- cor(data2)

# Create heatmaps with clustering and capture them as grobs
set.seed(123)
heatmap1 <- pheatmap(cor1, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F, 
                     main = "Healthy-weight", 
                     silent = TRUE)

heatmap2 <- pheatmap(cor2, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F,
                     main = "Obese",
                     silent = TRUE)

# p1 <- heatmap1$gtable # no need for these 2 anymore but keeping it just in case.
# p2 <- heatmap2$gtable

# Convert pheatmap objects to ggplot objects
p1_gg <- ggplotify::as.ggplot(heatmap1)
p2_gg <- ggplotify::as.ggplot(heatmap2)


# Apply the function to both plots
p1_gg <- set_square_aspect_ratio(p1_gg)
p2_gg <- set_square_aspect_ratio(p2_gg)

# Arrange the plots side by side using ggarrange
ggarrange(p1_gg, p2_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))

```

### BMI and Meds 

Make sure to note the imbalanced sample size between those who took vs did not take lipid meds. 

#### Hw taking Lipid meds vs Hw not taking Lipid meds
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load datasets
data3 <- healthy_nodm_NOmeds_hw %>% select(1:895)  
data4 <- healthy_nodm_meds_hw %>% select(1:895) 

# Calculate the correlation matrices
cor3 <- cor(data3)
cor4 <- cor(data4)

# Create heatmaps with clustering and capture them as grobs
set.seed(123)
heatmap3 <- pheatmap(cor3, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F, 
                     main = "Hw NO Lipid Med", 
                     silent = TRUE)

heatmap4 <- pheatmap(cor4, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F,
                     main = "Hw YES Lipid Med",
                     silent = TRUE)


# Convert pheatmap objects to ggplot objects
p3_gg <- ggplotify::as.ggplot(heatmap3)
p4_gg <- ggplotify::as.ggplot(heatmap4)


# Apply the function to both plots
p3_gg <- set_square_aspect_ratio(p3_gg)
p4_gg <- set_square_aspect_ratio(p4_gg)

# Arrange the plots side by side using ggarrange
ggarrange(p3_gg, p4_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))

```


#### Ob taking Lipid meds vs Ob not taking Lipid meds
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load datasets
data5 <- healthy_nodm_NOmeds_ob %>% select(1:895)  
data6 <- healthy_nodm_meds_ob %>% select(1:895) 

# Calculate the correlation matrices
cor5 <- cor(data5)
cor6 <- cor(data6)

# Create heatmaps with clustering and capture them as grobs
set.seed(123)
heatmap5 <- pheatmap(cor5, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F, 
                     main = "Ob NO Lipid Med", 
                     silent = TRUE)

heatmap6 <- pheatmap(cor6, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F,
                     main = "Ob YES Lipid Med",
                     silent = TRUE)

# Convert pheatmap objects to ggplot objects
p5_gg <- ggplotify::as.ggplot(heatmap5)
p6_gg <- ggplotify::as.ggplot(heatmap6)


# Apply the function to both plots
p5_gg <- set_square_aspect_ratio(p5_gg)
p6_gg <- set_square_aspect_ratio(p6_gg)

# Arrange the plots side by side using ggarrange
ggarrange(p5_gg, p6_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))

```


#### Hw NOT taking Lipid meds vs Ob not taking Lipid meds
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Arrange the heatmaps side by side
ggarrange(p3_gg, p5_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))
```

#### Hw taking Lipid meds vs Ob taking Lipid meds
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Arrange the heatmaps side by side
ggarrange(p4_gg, p6_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))
```



## Feature Selection 
### Multivariate modeling

Applying penalised regression models to lipid data (in this instance Healthy is simply No DM). 

```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(glmnet)
library(caret)

# Setting the data 
set.seed(123)
X <- healthy_nodm[, 1:895]
y <- healthy_nodm$bmi

# Split the data into training and testing sets
trainIndex <- createDataPartition(y, p = 0.8, list = FALSE)
X_train <- X[trainIndex, ]
y_train <- y[trainIndex]
X_test <- X[-trainIndex, ]
y_test <- y[-trainIndex]

# Convert data frames to matrices
X_train_matrix <- scale(as.matrix(X_train))
X_test_matrix <- scale(as.matrix(X_test), center = attr(X_train_matrix, "scaled:center"), scale = attr(X_train_matrix, "scaled:scale"))
```


#### Lasso
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Fit Lasso regression
lasso_model <- cv.glmnet(X_train_matrix, y_train, alpha = 1)

# Best lambda for Lasso
best_lambda_lasso <- lasso_model$lambda.min

# Extract coefficients at the best lambda
lasso_coefs <- coef(lasso_model, s = best_lambda_lasso)

# Convert the sparse matrix to a readable format
lasso_coefs_df <- as.data.frame(as.matrix(lasso_coefs))

# Display non-zero coefficients
lasso_nonzero_coefs <- lasso_coefs_df[lasso_coefs_df != 0, , drop = FALSE]
#print(lasso_nonzero_coefs)

# Predict on the test set
lasso_predictions <- predict(lasso_model, s = best_lambda_lasso, newx = X_test_matrix)

```

#### Ridge
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Fit Ridge regression
ridge_model <- cv.glmnet(X_train_matrix, y_train, alpha = 0)

# Best lambda for Ridge
best_lambda_ridge <- ridge_model$lambda.min

# Extract coefficients at the best lambda
ridge_coefs <- coef(ridge_model, s = best_lambda_ridge)

# Convert the sparse matrix to a readable format
ridge_coefs_df <- as.data.frame(as.matrix(ridge_coefs))

# Display all coefficients
#print(ridge_coefs_df)


# Predict on the test set
ridge_predictions <- predict(ridge_model, s = best_lambda_ridge, newx = X_test_matrix)
```


#### Elastic net
```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Fit Elastic Net model
elastic_net_model <- cv.glmnet(X_train_matrix, y_train, alpha = 0.5)  # alpha = 0.5 for equal balance

# Best lambda for Elastic Net
best_lambda_elastic_net <- elastic_net_model$lambda.min


# Extract coefficients at the best lambda
elastic_net_coefs <- coef(elastic_net_model, s = best_lambda_elastic_net)

# Convert the sparse matrix to a readable format
elastic_net_coefs_df <- as.data.frame(as.matrix(elastic_net_coefs))

# Display non-zero coefficients
elastic_net_nonzero_coefs <- elastic_net_coefs_df[elastic_net_coefs_df != 0, , drop = FALSE]
#print(elastic_net_nonzero_coefs)


# Predict on the test set
elastic_net_predictions <- predict(elastic_net_model, s = best_lambda_elastic_net, newx = X_test_matrix)

```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Identify non-zero coefficients for Lasso and Elastic Net
lasso_nonzero_predictors <- rownames(lasso_coefs_df)[lasso_coefs_df[,1] != 0]
elastic_net_nonzero_predictors <- rownames(elastic_net_coefs_df)[elastic_net_coefs_df[,1] != 0]

# For Ridge, consider non-zero coefficients (they are typically non-zero due to regularization)
ridge_nonzero_predictors <- rownames(ridge_coefs_df)[ridge_coefs_df[,1] != 0]

# Find common predictors identified by Lasso, Ridge, and Elastic Net
common_predictors <- Reduce(intersect, list(lasso_nonzero_predictors, elastic_net_nonzero_predictors, ridge_nonzero_predictors))

print("Common Predictors Identified by All Models:")
print(common_predictors)

```


#### Accuracy measurement
```{r, echo=FALSE, warning=FALSE, message=FALSE}
accuracy_metrics <- function(true_values, predicted_values) {
  mse <- mean((true_values - predicted_values)^2)
  rmse <- sqrt(mse)
  r2 <- 1 - (sum((true_values - predicted_values)^2) / sum((true_values - mean(true_values))^2))
  return(list(MSE = mse, RMSE = rmse, R2 = r2))
}
```


```{r, echo=FALSE, warning=FALSE, message=FALSE}
# Calculate accuracy for Lasso
lasso_accuracy <- accuracy_metrics(y_test, lasso_predictions)
# print("Lasso Regression Accuracy:")
# print(lasso_accuracy)

# Calculate accuracy for Ridge
ridge_accuracy <- accuracy_metrics(y_test, ridge_predictions)
# print("Ridge Regression Accuracy:")
# print(ridge_accuracy)

# Calculate accuracy for Elastic Net
elastic_net_accuracy <- accuracy_metrics(y_test, elastic_net_predictions)
# print("Elastic Net Regression Accuracy:")
# print(elastic_net_accuracy)

model_performance_summary <- data.frame("Model" = c("Lasso", "Ridge", "Elastic Net"), "MSE" = c(lasso_accuracy$MSE, ridge_accuracy$MSE, elastic_net_accuracy$MSE), "RMSE" = c(lasso_accuracy$RMSE, ridge_accuracy$RMSE, elastic_net_accuracy$RMSE), "R2" = c(lasso_accuracy$R2, ridge_accuracy$R2, elastic_net_accuracy$R2))

print("Summary of Model Performance: ")
print(model_performance_summary)

```



# Split 2: selecting Healthy and Obese (but not diabetes), T2DM groups
Healthy = Summary_Healthy==1 means No conditions AND Healthy weight. 
```{r, echo=FALSE, message= FALSE, warning=FALSE}
idx_hty = which(combined$Summary_Healthy==1) #255 #No conditions AND Healthy weight. 
idx_obs = which(combined$Summary_overweight_idx==1 & combined$Summary_Diabetes_idx==0) #466 #Overweight and Obese and Not Diabetic
idx_DM = which(combined$Summary_Diabetes_idx==1) #160 # Diabetic regardless BMI


Combined_subData<-combined[c(idx_hty,idx_obs,idx_DM),] 

Combined_subData <- Combined_subData%>%
  mutate(Group = ifelse(Summary_Healthy==1,"Hty",NA),
         Group = ifelse(Summary_overweight_idx==1,"OBS",Group),  #overweight & obese
         Group = ifelse(Summary_Diabetes_idx==1,"DM",Group))%>%
  relocate(Group,.after = bloodno6) %>% relocate(sampleID,.before = "TPTG") 


Combined_subData %>% select(Group,sex,age,bmi,cholesterol,triglycerides,high_density_lipoprotein,glucose,c_reactive_protein,creatinine,glycated_haemoglobin,sbp,dbp,Summary_Hypertension_idx:Summary_Diabetes_idx, taking_lipid_med, taking_any_med)  %>% gtsummary::tbl_summary(by = Group)

#idx_hty_taking_meds <- which(Combined_subData$taking_any_med == 1 & Combined_subData$Summary_Healthy ==1) #35 who are deemed super healthy but taking some sort of meds - BUT NOT taking lipid med.

#Combined_subData[idx_hty_taking_meds, ] %>% select(BMI.weight, an_med) 

Hw <- Combined_subData[idx_hty, ] #255
Ob <- Combined_subData[idx_obs, ] #466 #includes the overweights

```


## Heatmap with Clustering
Further split by Sex - M_Hw v F_Hw; M_Ob v F_Ob

### Male Hw v Ob
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load datasets
data7a <- Hw %>% filter(Hw$sex=="M") %>% select(1:895) #50   
data7b <- Ob %>% filter(Ob$sex=="M")%>% select(1:895) #106

# Calculate the correlation matrices
cor7a <- cor(data7a)
cor7b <- cor(data7b)

# Create heatmaps with clustering and capture them as grobs
set.seed(123)
heatmap7a <- pheatmap(cor7a, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F, 
                     main = "Male Hw", 
                     silent = TRUE)

heatmap7b <- pheatmap(cor7b, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F,
                     main = "Male Obese",
                     silent = TRUE)

# Convert pheatmap objects to ggplot objects
p7a_gg <- ggplotify::as.ggplot(heatmap7a)
p7b_gg <- ggplotify::as.ggplot(heatmap7b)


# Apply the function to both plots
p7a_gg <- set_square_aspect_ratio(p7a_gg)
p7b_gg <- set_square_aspect_ratio(p7b_gg)

# Arrange the plots side by side using ggarrange
ggarrange(p7a_gg, p7b_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))

```



### Female Hw v Ob
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load datasets
data8a <- Hw %>% filter(Hw$sex=="F") %>% select(1:895) #70   
data8b <- Ob %>% filter(Ob$sex=="F")%>% select(1:895) #110

# Calculate the correlation matrices
cor8a <- cor(data8a)
cor8b <- cor(data8b)

# Create heatmaps with clustering and capture them as grobs
set.seed(123)
heatmap8a <- pheatmap(cor8a, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F, 
                     main = "Female Hw", 
                     silent = TRUE)

heatmap8b <- pheatmap(cor8b, 
                     clustering_method = "complete", show_rownames = F, show_colnames = F,
                     main = "Female Obese",
                     silent = TRUE)

# Convert pheatmap objects to ggplot objects
p8a_gg <- ggplotify::as.ggplot(heatmap8a)
p8b_gg <- ggplotify::as.ggplot(heatmap8b)


# Apply the function to both plots
p8a_gg <- set_square_aspect_ratio(p8a_gg)
p8b_gg <- set_square_aspect_ratio(p8b_gg)

# Arrange the plots side by side using ggarrange
ggarrange(p8a_gg, p8b_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))

```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggarrange(p7a_gg, p8a_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))
ggarrange(p7b_gg, p8b_gg, ncol = 2, nrow = 1, widths = c(1, 1), heights = c(1, 1))
```


## Feature Selection 
### Lipid Summary Table 
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
### Lipids summary table

Combined_subData%>%
  dplyr::select(Group,lipid_names)%>%
      group_by(Group) %>%
      summarise(across(where(is.numeric), 
                       list(mean = ~mean(., na.rm = TRUE),
                            sd = ~sd(., na.rm = TRUE)), 
                       .names = "{.col}__{.fn}")) %>%
      pivot_longer(cols = !c(Group), names_to = "Variable", values_to = "Value")%>%
      separate(Variable, into = c("Variable", "Statistic"), sep = "__")%>%
      pivot_wider(names_from = "Statistic",values_from = "Value")%>%
      mutate(values = paste0(round(mean,2)," ± ",round(sd,2)))%>%
      select(!mean:sd)%>%
      pivot_wider(names_from = Group,values_from = values)->tbl1


tbl1<-perform_stat_tests(Combined_subData,group = "Group",var = lipid_names,method = "anova",tbl1)

```


## Univariate Modelling 

In the context of the univariate t-test results generated using the broom::tidy() function, the terms estimate, estimate1, estimate2 and parameter refer to specific statistical values related to the groups being compared: 

- estimate: This is the estimated difference in means between the two groups. It represents the difference in the average values of the dependent variable for the two levels of the grouping variable.

- estimate1: This is the mean value of the dependent variable for the first group (the group that appears first when sorted alphabetically or numerically).

- estimate2: This is the mean value of the dependent variable for the second group (the group that appears second when sorted alphabetically or numerically).

- parameter: Typically refers to the degrees of freedom (df). 


Bonferroni correction has been applied and P-value threshold has been adjusted to account for multiple testing. 

Only lipids that have p-value < 0.0001 considered as significant in this instance.

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# creating obese01 variable for t-test 
Combined_subData <- Combined_subData %>% mutate(obese01 = ifelse(BMI.weight == "Obese", 1, 0))
lipid_subData <- Combined_subData %>% select(1:896, "sex", "obese01") 

# Apply the univariate test function to all variables - regardless sex
results <- lipid_subData %>%
  pivot_longer(cols = -c(sampleID, obese01, sex), names_to = "variable", values_to = "value") %>%
  mutate(group_var = obese01) %>%
  group_by(variable) %>%
  nest() %>%
  mutate(test_result = map(data, univariate_test)) %>%
  unnest(test_result) %>%
  select(-data)

# Adjusting for multiple testing
results$adjusted_p_value <- p.adjust(results$p.value, method = "bonferroni")

# Filter significant results based on adjusted p-value threshold
significant_results <- results %>%
  filter(adjusted_p_value < 0.0001) %>% arrange(adjusted_p_value) # 235 lipids- still quite a lot
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Apply the function to all variables - M v F separately
results_M <- lipid_subData %>% filter(sex == "M") %>% 
  pivot_longer(cols = -c(sampleID, obese01, sex), names_to = "variable", values_to = "value") %>%
  mutate(group_var = obese01) %>%
  group_by(variable) %>%
  nest() %>%
  mutate(test_result = map(data, univariate_test)) %>%
  unnest(test_result) %>%
  select(-data)

# Adjusting for multiple testing
results_M$adjusted_p_value <- p.adjust(results_M$p.value, method = "bonferroni")

# Filter significant results based on adjusted p-value threshold
significant_results_M <- results_M %>%
  filter(adjusted_p_value < 0.0001) %>% arrange(adjusted_p_value) 

# p-value = 0.05 -> 235 sig 
# p-value = 0.01 -> 132 sig
# p-value = 0.001 -> 84 sig
# p-value = 0.0001 ->  47 sig

print("Top 20 (out of 47) significant Lipids in Male:")
print(head(significant_results_M, 20))

```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Apply the function to all variables - M v F separately
results_F <- lipid_subData %>% filter(sex == "F") %>% 
  pivot_longer(cols = -c(sampleID, obese01, sex), names_to = "variable", values_to = "value") %>%
  mutate(group_var = obese01) %>%
  group_by(variable) %>%
  nest() %>%
  mutate(test_result = map(data, univariate_test)) %>%
  unnest(test_result) %>%
  select(-data)

# Adjusting for multiple testing
results_F$adjusted_p_value <- p.adjust(results_F$p.value, method = "bonferroni")

# Filter significant results based on adjusted p-value threshold 
significant_results_F <- results_F %>%
  filter(adjusted_p_value < 0.0001) %>% arrange(adjusted_p_value) 

print("Top 20 (out of 147) significant Lipids in Female:")
print(head(significant_results_F, 20))

# p-value = 0.05 -> 312 sig 
# p-value = 0.01 -> 270 sig
# p-value = 0.001 ->  201 sig
# p-value = 0.0001 ->  147 sig # almost 3 times as Male ones. 

common_sig_lipids <- Reduce(intersect, list(significant_results_M$variable, significant_results_F$variable)) # 40 lipids sig in both M & F
print("Lipids significant in Male and Female:")
print(common_sig_lipids) 
lipids_in_M_only <- setdiff(significant_results_M$variable, significant_results_F$variable) # 7 lipids sig in M only
print("Lipids significant in Male only:")
print(lipids_in_M_only)
lipids_in_F_only <- setdiff(significant_results_F$variable, significant_results_M$variable) # 107 lipids sig in F only
print("Lipids significant in Female only:")
print(lipids_in_F_only)

```

```{r}
# if (!require("BiocManager", quietly = TRUE))
#        install.packages("BiocManager")
# BiocManager::install(c("GO.db", "preprocessCore", "impute"), force = TRUE )
#install.packages("WGCNA")   # WGCNA is available on CRAN

# Restart R session to install DBI package
# .rs.restartR()
# install.packages("DBI", repos = "http://cran.us.r-project.org")

# Load the WGCNA package
library(WGCNA)

```

# WGCNA
## Data preparation
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load your metabolomics data
# must be a data frame where rows represent samples and columns represent lipid
data <-lipid_subData[, 1:895]

# Ensure the data is numeric - my lipid data is already numeric so no need for this step
#data <- as.data.frame(lapply(data, as.numeric))

# Check for missing values
gsg <- goodSamplesGenes(data, verbose = 3)
if (!gsg$allOK) {
  # If some samples/lipid have too many missing values, consider removing them
  data <- data[gsg$goodSamples, gsg$goodGenes]
}
```

## Choose a Soft-Thresholding Power
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Choose a set of soft-thresholding powers
powers <- c(1:20)
sft <- pickSoftThreshold(data, powerVector = powers, verbose = 5)

# Plot the results to choose the best power
sizeGrWindow(9, 5)
par(mfrow = c(1, 2))
cex1 <- 0.9
plot(sft$fitIndices[, 1], -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     type = "n", main = "Scale independence")
text(sft$fitIndices[, 1], -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     labels = powers, cex = cex1, col = "red")
abline(h = 0.90, col = "red")  # This line corresponds to using an R^2 cut-off of h

plot(sft$fitIndices[, 1], sft$fitIndices[, 5],
     xlab = "Soft Threshold (power)", ylab = "Mean Connectivity", type = "n",
     main = "Mean connectivity")
text(sft$fitIndices[, 1], sft$fitIndices[, 5], labels = powers, cex = cex1, col = "red")
```

## Construct the Network and Identify Modules
```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Choose the power based on the above plot
softPower <- 8 

# Construct the adjacency matrix
adjacency <- adjacency(data, power = softPower)

# Turn adjacency into topological overlap matrix (TOM) and calculate the dissimilarity
TOM <- TOMsimilarity(adjacency)
dissTOM <- 1 - TOM

# Perform hierarchical clustering
geneTree <- hclust(as.dist(dissTOM), method = "average")

# Module identification using dynamic tree cut
dynamicMods <- cutreeDynamic(dendro = geneTree, distM = dissTOM,
                             deepSplit = 2, pamRespectsDendro = FALSE,
                             minClusterSize = 30)

# Convert numeric labels into colors
dynamicColors <- labels2colors(dynamicMods)
table(dynamicColors)

# Plot the dendrogram and the module colors
sizeGrWindow(8, 6)
plotDendroAndColors(geneTree, dynamicColors, "Dynamic Tree Cut",
                    dendroLabels = FALSE, hang = 0.03,
                    addGuide = FALSE, guideHang = 0.05)
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
# Load your outcome/trait data
traits <- data.frame("obese" = lipid_subData[,898])

# Match samples between metabolomics data and trait data - lipid and obesity data are matched already
# matchedSamples <- intersect(rownames(data), rownames(traits))
# data <- data[matchedSamples, ]
# traits <- traits[matchedSamples, ]

# Calculate module eigengenes
MEs <- moduleEigengenes(data, colors = dynamicColors)$eigengenes
MEs <- orderMEs(MEs)

# Correlate module eigengenes with external traits
moduleTraitCor <- cor(MEs, traits, use = "p")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nrow(data))

# Plot the heatmap of module-trait relationships
sizeGrWindow(10, 6)
textMatrix <- paste(signif(moduleTraitCor, 2), "\n(",
                    signif(moduleTraitPvalue, 1), ")", sep = "")
dim(textMatrix) <- dim(moduleTraitCor)
par(mar = c(6, 8.5, 3, 3))
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = names(traits),
               yLabels = names(MEs),
               ySymbols = names(MEs),
               colorLabels = FALSE,
               colors = blueWhiteRed(50),
               textMatrix = textMatrix,
               setStdMargins = FALSE,
               cex.text = 0.5,
               zlim = c(-1, 1),
               main = paste("Module-trait relationships"))


```

## Network heatmap plot visulization
```{r}
# Calculate topological overlap anew: this could be done more efficiently by saving the TOM
# calculated during module detection, but let us do it again here.
dissTOM = 1-TOMsimilarityFromExpr(data, power = softPower)
# Transform dissTOM with a power to make moderately strong connections more visible in the heatmap
plotTOM = dissTOM^3;
# Set diagonal to NA for a nicer plot
diag(plotTOM) = NA
sizeGrWindow(9,9)
TOMplot(plotTOM, geneTree, dynamicColors, main = "Network heatmap plot")

```


