---
title: "Busselton Lipo-Lipid"
author: "nminaee"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
```

```{r}
library(dplyr)
library(readr)
library(reshape2)
library(ggplot2)
library(ggrepel)
library(tibble)
library(ggpubr)
library(mva.plots)
library(ca)
library(tidyr)
library(tidyverse)
library(corrplot)
library(GGally)
library(stats)
library(pheatmap)
library(cluster)
library(gridExtra)
library(ggpubr)
library(htmlTable)
library(broom)
```


```{r}
# Create a function to set square aspect ratio
set_square_aspect_ratio <- function(plot) {
  plot + theme(aspect.ratio = 1)
}


continuousPalette<- c(
                        "#0000CC",
                        "#0000FF",
                        "#0055FF",
                        "#00AAFF",
                        "#00FFFF",
                        "#2BFFD5",
                        "#55FFAA",
                        "#80FF80",
                        "#AAFF55",
                        "#D4FF2B",
                        "#FFFF00",
                        "#FFAA00",
                        "#FF5500",
                        "#FF0000",
                        "#CC0000")

txt2S <- function(col) {
  newCol <- lapply(col, function(i) {
    if (!is.na(i)){
      i <- abs(as.numeric(i))
      if (i < 0.05 & i >= 0.01) {"*"
      } else if (i < 0.01 & i >= 0.001) {
        "**"
      } else if (i < 0.001) {
        "***"
      } else {
        ""
      }
    } else {
      NA
    }
  })
  return(unlist(newCol))
}


# Function to perform univariate t-test
univariate_test <- function(data) {
  t.test(value ~ group_var, data = data) %>% 
    broom::tidy()
}



perform_stat_tests<-function(tdf_tbl,group = NULL,var, method, tbl1){
  # Handle missing or infinite values
  data_pval <- tdf_tbl %>%
    mutate(across(-{{group}}, ~replace(., is.nan(.) | is.infinite(.), 0)))%>%
    rename(group = {{group}})%>%
    dplyr::select(group,var)
  
  
  # Switch between ANOVA and Wilcoxon tests based on the method argument
  p_model <- switch(method,
                    "anova" = lapply(data_pval %>% select(-group), function(x) {
                      aov(x ~ data_pval$group)
                    }),
                    "wilcox" = {
                      if (length(unique(data_pval$group)) > 2) {
                        stop("Wilcoxon test does not support more than 2 groups naturally.")
                      } else {
                        lapply(data_pval %>% select(-group), function(x) {
                          wilcox.test(x ~ data_pval$group)
                        })
                      }
                    },
                    stop("Unsupported method type provided")
  )
  
  # Generate summaries of the model results
  model_summaries <- lapply(p_model, summary)
  
  
  # Extracting results and performing post-hoc tests where applicable
  summary_list <- list()
  
  for(i in names(model_summaries)){
    variable = data.frame(Variable =i)
    variable$Pvalue = txt2S(model_summaries[[i]][[1]]['Pr(>F)'][1,])
    tukey_p<-as.data.frame(TukeyHSD(p_model[[i]])[[1]])
    tukey_p$comparison<-rownames(tukey_p)
    tukey_p$`p adj`<-txt2S(tukey_p$`p adj`)
    tukey_p<-data.frame(t(tukey_p[,c("comparison","p adj")]))
    colnames(tukey_p)<-tukey_p[1,]
    tukey_p<-tukey_p[-1,]
    variable<-cbind(variable,tukey_p)
    summary_list[[i]]<-variable
    rm(variable,tukey_p)
  }
  
  # Combine all individual summaries into a single data frame
  summary_list <- do.call(rbind, summary_list)
  
  # Merge the summarized data frame with the existing table `tbl1`
  tbl1 <- merge(tbl1, summary_list, by = "Variable")
  
  return(tbl1)
}
```

## Load the data

Note: I only use Busselton's timepoint 1 and T2DM group for diabetes (T1DM is excluded). 

### Lipo

Metadata: Anno_t1_MH_new.csv

Lipoprotein data: Lipo_bus_230724.csv 

(both from Sam L)
```{r}
#loading buss metadata for Timepoint 1 from Sam 
Annotation <- read.csv("/Users/novia/Desktop/Diabetes_Obesity/Busselton/Anno_t1_MH_new.csv", header=TRUE) # t1 = timepoint 1

#loading latest buss lipo data from Sam (23/07/2024)
Lipo<- read.csv(file='/Users/novia/Desktop/Diabetes_Obesity/Busselton/Lipo_bus_230724.csv') #3945 obs - T1 and T2 but I'll only use T1

# matching metadata and lipoprotein 
idx<-which(Annotation$sampleID %in% Lipo$sampleID) # 1976
Annotation<-Annotation[which(Annotation$sampleID %in% Lipo$sampleID),]

idx<-which(Lipo$sampleID %in% Annotation$sampleID)  #1976
Lipo<-Lipo[idx,]

Annotation<-Annotation[match(Lipo$sampleID, Annotation$sampleID),]


#rm(idx)
```


### Lipids

Metadata: the same as Lipo above

Lipid data: Busselton_PLA_MS_Lipids.daE

```{r}
da <- local(get(load("/Users/novia/Desktop/Diabetes_Obesity/Busselton/LipidMS/Busselton_PLA_MS_Lipids.daE")))
Lipids<-data.frame(apply(da@.Data,2,as.numeric))
Lipids_ANN<-da@obsDescr[[1]]
idx<-match(Annotation$sampleID,Lipids_ANN$sampleID)

Lipids<-Lipids[which(Lipids_ANN$sampleID %in% Annotation$sampleID),] #1976 
Lipids_ANN<-Lipids_ANN[which(Lipids_ANN$sampleID %in% Annotation$sampleID),] #1976 but use the Busselton's metadata (Annotation), not the Lipids_Ann for further analysis. 

#rm(idx, da)
```


```{r}
lipo_names<-colnames(Lipo[, 2:113]) # only the lipo, minus sampleID.)
lipid_names<-colnames(Lipids)
lipid_class<-unique(sapply(strsplit(lipid_names,".",fixed = T),"[",1))
```


## Three main groups: Healthy, Obese, DM

- Healthy
  
  - no health issue or condition (Summary_Healthy = 1)
  - healthy weight (BMI < 30) 
  
    This is slightly more generous than the WHO BMI indexing but since this is an older cohort, we allow a higher BMI range for the healthy weight. 
    WHO BMI indexing would include those 25 - 30  as overweight - this is an important fact to note when comparing to another cohort.
    

- Obese 
  
  - not diabetic (Summary_Type2_Diabetes_idx = 0)
  - can have other health conditions
  - BMI > 30 
  
- T2DM 

  - has Type 2 diabetes (Summary_Type2_Diabetes_idx = 1)
  - can have other health conditions
  - BMI > 27.7 (erring more towards the obese side)



## Background summary table
```{r}
idx_hty = which(Annotation$Summary_Healthy==1)
idx_obs = which(Annotation$Summary_overweight_idx==1 & Annotation$Summary_Diabetes_idx==0)
idx_DM = which(Annotation$Summary_Diabetes_idx==1)

Combined_Data<-data.frame(Annotation,Lipo,Lipids)
Combined_Data$BMI_over30 <- ifelse(Combined_Data$bmi>30, 1,0) 

idx_DM_H = which(Annotation$Summary_Diabetes_idx==1 & Combined_Data$BMI_over30 ==0)
idx_DM_O = which(Annotation$Summary_Diabetes_idx==1 & Combined_Data$BMI_over30 ==1)

#Combined_subData<-Combined_Data[c(idx_hty,idx_obs,idx_DM),]
Combined_subData <- Combined_Data[c(idx_hty,idx_obs, idx_DM_H, idx_DM_O),]


Combined_subData<-Combined_subData%>%
  mutate(Group = ifelse(Summary_Healthy==1,"Healthy",NA),
         Group = ifelse(Summary_overweight_idx==1,"Obese",Group),
         Group = ifelse(Summary_Diabetes_idx==1,"T2DM",Group))%>%
  relocate(Group,.after = bloodno6)


Combined_subData%>%
  select(Group,sex,age,bmi,cholesterol,triglycerides,high_density_lipoprotein,glucose,c_reactive_protein,Summary_Hyperlipidemia_idx:Summary_Diabetes_idx)%>%
  gtsummary::tbl_summary(by = Group, missing = "no")

#rm(idx_DM, idx_hty, idx_obs)
```


## Lipoprotein summary table

P-value = Anova result

- if significant, post-hoc tests amongst the groups (Tukey's HSD) are performed to identify which specific groups differ from each other. 

</div>
```{r}
Combined_subData%>%
  dplyr::select(Group,lipo_names)%>%
      group_by(Group) %>%
      summarise(across(where(is.numeric), 
                       list(mean = ~mean(., na.rm = TRUE),
                            sd = ~sd(., na.rm = TRUE)), 
                       .names = "{.col}_{.fn}")) %>%
      pivot_longer(cols = !Group, names_to = "Variable", values_to = "Value")%>%
      separate(Variable, into = c("Variable", "Statistic"), sep = "_")%>%
      pivot_wider(names_from = "Statistic",values_from = "Value")%>%
      mutate(values = paste0(round(mean,2)," ± ",round(sd,2)))%>%
      select(!mean:sd)%>%
      pivot_wider(names_from = Group,values_from = values)->tbl1
tbl1<-perform_stat_tests(Combined_subData,group = "Group",var = lipo_names,method = "anova",tbl1)

htmlTable(tbl1, css.table = "width: 100%;") # make the columns more spaced out

#write.csv(tbl1, "/Users/novia/Desktop/Diabetes_Obesity/Busselton/Lipo_summary_table_23072024.csv", row.names = F)
```

<div style="margin-bottom:50px;">


</div>
```{r}
# lipos that are significant only in 1 group comparisons 
tbl1_1p <- tbl1 %>%
  filter(Pvalue != "" &
         rowSums(across(c("Obese-Healthy", "T2DM-Healthy", "T2DM-Obese"), ~ . == "***" | . == "**" | . == "*")) < 2)

htmlTable(tbl1_1p, caption = "Table 1: Lipoproteins that are significant only in 1 group comparison, n = 7",css.table = "width: 100%;") # 7 lipos

```
<div style="margin-bottom:50px;">

</div>
```{r}
# lipos that are significant in 2 group comparisons 
tbl1_2p <- tbl1 %>%
  filter(Pvalue != "" &
         rowSums(across(c("Obese-Healthy", "T2DM-Healthy", "T2DM-Obese"), ~ . == "***" | . == "**" | . == "*")) == 2)

htmlTable(tbl1_2p, caption = "Table 2: Lipoproteins that are significant in 2 group comparisons, n = 77", css.table = "width: 100%;") # 77 lipos

```
<div style="margin-bottom:50px;">



</div>
```{r}
# lipos that are significant on all 3 group comparisons 
tbl1_3p <- tbl1 %>%
  filter(Pvalue != "" &
         rowSums(across(c("Obese-Healthy", "T2DM-Healthy", "T2DM-Obese"), ~ . == "***" | . == "**" | . == "*")) == 3)

htmlTable(tbl1_3p, caption = "Table 3: Lipoproteins that are significant in all 3 group comparisons, n = 23", css.table = "width: 100%;") # 23 lipos
```
<div style="margin-bottom:50px;">



</div>
```{r}
# lipos that are NOT significant at all
tbl1_ns <- tbl1 %>%
  filter(Pvalue == "")

htmlTable(tbl1_ns, caption = "Table 4: Lipoproteins that are NOT significant, n = 5", css.table = "width: 100%;") # 5 lipos


```
<div style="margin-bottom:50px;">


<br>

#### 77 lipoproteins are found to be significant in at least 2 pairs of group comparison. They will be used for further downstream analysis. 


## Correlation analysis 

- To figure out the relationship between the significant lipoproteins and the lipids. 

- The analysis will be done per group. 


```{r}
lipo_2p<-unique(colnames(Combined_subData[colnames(Combined_subData) %in% tbl1_2p$Variable]))

subLipo <- Combined_subData %>%
  select(Group, sex, age_cat, all_of(lipo_2p))

subLipoH <- subLipo[idx_hty, ]
subLipoO <- subLipo[idx_obs, ]
subLipoD <- subLipo[idx_DM, ]

subLipid <- Combined_subData %>% 
select(Group, sex, age_cat, "CE.14.0.":ncol(Combined_subData))

subLipidH <- subLipid[idx_hty, ]
subLipidO <- subLipid[idx_obs, ]
subLipidD <- subLipid[idx_DM, ]


```


```{r}
# Extract the lipo and lipid from Healthy subLipo dataset
# data1 and data2 are already dataframes. 
data1 <- subLipoH %>% select(4:ncol(subLipoH)) 
data2 <- subLipidH %>% select(4:ncol(subLipidH)) 

# Initialize a matrix to store the correlation coefficients  - m rows (lipo) x n columns (lipids)
cor_matrixH <- matrix(NA, nrow = ncol(data1), ncol = ncol(data2))
rownames(cor_matrixH) <- colnames(data1)
colnames(cor_matrixH) <- colnames(data2)

# Compute the correlations
for(i in 1:ncol(data1)) {
  for(j in 1:ncol(data2)) {
    cor_matrixH[i, j] <- cor(data1[[i]], data2[[j]], use = "complete.obs")
  }
}

# Convert the matrix to a data frame for easier handling
cor_dfH <- as.data.frame(cor_matrixH)

# Generate the heatmap
pH <- pheatmap(cor_dfH, cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = FALSE, show_colnames = FALSE, silent = TRUE, main = "Lipo-Lipid Healthy")

print(pH)
```



```{r}
# Extract the lipo and lipid from Obese subLipo dataset
# data3 and data4 are already dataframes. 
data3 <- subLipoO %>% select(4:ncol(subLipoO)) 
data4 <- subLipidO %>% select(4:ncol(subLipidO)) 

# Initialize a matrix to store the correlation coefficients  - m rows (lipo) x n columns (lipids)
cor_matrixO <- matrix(NA, nrow = ncol(data3), ncol = ncol(data4))
rownames(cor_matrixO) <- colnames(data3)
colnames(cor_matrixO) <- colnames(data4)

# Compute the correlations
for(i in 1:ncol(data3)) {
  for(j in 1:ncol(data4)) {
    cor_matrixO[i, j] <- cor(data3[[i]], data4[[j]], use = "complete.obs")
  }
}

# Convert the matrix to a data frame for easier handling
cor_dfO <- as.data.frame(cor_matrixO)

# Generate the heatmap
pO <- pheatmap(cor_dfO, cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = FALSE, show_colnames = FALSE, silent = TRUE, main = "Lipo-Lipid Obese")

print(pO)
```


```{r}
# Extract the lipo and lipid from T2DM subLipo dataset
#data5 and data6 are already dataframes.
data5 <- subLipoD %>% select(4:ncol(subLipoD)) 
data6 <- subLipidD %>% select(4:ncol(subLipidD))

# Initialize a matrix to store the correlation coefficients - m rows (lipo) x n columns (lipids)
cor_matrixD <- matrix(NA, nrow = ncol(data5), ncol = ncol(data6))
rownames(cor_matrixD) <- colnames(data5)
colnames(cor_matrixD) <- colnames(data6)

# Compute the correlations
for(i in 1:ncol(data5)) {
  for(j in 1:ncol(data6)) {
    cor_matrixD[i, j] <- cor(data5[[i]], data6[[j]], use = "complete.obs")
  }
}

# Convert the matrix to a data frame for easier handling
cor_dfD <- as.data.frame(cor_matrixD)

# Generate the heatmap
pD <- pheatmap(cor_dfD, cluster_rows = TRUE, cluster_cols = TRUE, show_rownames = FALSE, show_colnames = FALSE, silent = TRUE, main = "Lipo - Lipid T2DM")

print(pD)
```



## PCA of the correlation results

Conducting a Principal Component Analysis on the above correlation results between lipo and lipid. 

```{r}
library(plotly)

mod<-PCA(cor_dfH, plot=FALSE)

data.frame(mod$data$scores)%>%
  mutate(label = lipo_2p,
         class = substr(label,3,4),
         class = ifelse(label %in% c("ABA1","LDHD"),"ratio",class))%>%
  ggplot(aes(x = PC1,y = PC2, color = class, label = label))+
  geom_point()+geom_text()+
  labs(title = "PCA scores of Lipo-Lipid correlation for Healthy Group", col = "Lipo Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p1

data.frame(mod$data$loadings)%>%
  mutate(lipid_names = lipid_names,
         lipid_class = sapply(strsplit(lipid_names,".",fixed = TRUE),"[",1))%>%
  ggplot(aes(PC1,PC2, color = factor(lipid_class),label = lipid_names))+
  geom_point()+
  labs(title = "PCA Loadings of Lipo-Lipid correlation for Healthy Group", col = "Lipid Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p2


ggplotly(p1)

ggplotly(p2)

```


```{r}
mod<-PCA(cor_dfO, plot=FALSE)

data.frame(mod$data$scores)%>%
  mutate(label = lipo_2p,
         class = substr(label,3,4),
         class = ifelse(label %in% c("ABA1","LDHD"),"ratio",class))%>%
  ggplot(aes(x = PC1,y = PC2, color = class, label = label))+
  geom_point()+geom_text()+
  labs(title = "PCA scores of Lipo-Lipid correlation for Obese Group", col = "Lipo Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p3

data.frame(mod$data$loadings)%>%
  mutate(lipid_names = lipid_names,
         lipid_class = sapply(strsplit(lipid_names,".",fixed = TRUE),"[",1))%>%
  ggplot(aes(PC1,PC2, color = factor(lipid_class),label = lipid_names))+
  geom_point()+
  labs(title = "PCA Loadings of Lipo-Lipid correlation for Obese Group", col = "Lipid Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p4


ggplotly(p3)

ggplotly(p4)
```


```{r}
mod<-PCA(cor_dfD, plot=FALSE)

data.frame(mod$data$scores)%>%
  mutate(label = lipo_2p,
         class = substr(label,3,4),
         class = ifelse(label %in% c("ABA1","LDHD"),"ratio",class))%>%
  ggplot(aes(x = PC1,y = PC2, color = class, label = label))+
  geom_point()+geom_text()+
  labs(title = "PCA scores of Lipo-Lipid correlation for T2DM Group", col = "Lipo Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p5

data.frame(mod$data$loadings)%>%
  mutate(lipid_names = lipid_names,
         lipid_class = sapply(strsplit(lipid_names,".",fixed = TRUE),"[",1))%>%
  ggplot(aes(PC1,PC2, color = factor(lipid_class),label = lipid_names))+
  geom_point()+
  labs(title = "PCA Loadings of Lipo-Lipid correlation for T2DM Group", col = "Lipid Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p6


ggplotly(p5)

ggplotly(p6)
```

```{r}
mod<-PCA(cor_dfDH, plot=FALSE)

#plotScores(mod,optns = list(PCi = 1,PCj = 2))

data.frame(mod$data$scores)%>%
  mutate(label = lipo_4p,
         class = substr(label,3,4),
         class = ifelse(label %in% c("ABA1","LDHD"),"ratio",class))%>%
  ggplot(aes(x = PC1,y = PC2, color = class, label = label))+
  geom_point()+geom_text()+
  labs(title = "PCA scores of Lipo-Lipid correlation for T2DM_Hw Group", col = "Lipo Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p5

data.frame(mod$data$loadings)%>%
  mutate(lipid_names = lipid_names,
         lipid_class = sapply(strsplit(lipid_names,".",fixed = TRUE),"[",1))%>%
  ggplot(aes(PC1,PC2, color = factor(lipid_class),label = lipid_names))+
  geom_point()+
  labs(title = "PCA Loadings of Lipo-Lipid correlation for T2DM Group", col = "Lipid Class",
       x = paste0("PC1 (",round(mod$data$pcSum$`Proportion of Variance`[1],1),"%)"),
       y = paste0("PC2 (",round(mod$data$pcSum$`Proportion of Variance`[2],1),"%)"))+
  geom_hline(yintercept = 0,col = "grey")+
  geom_vline(xintercept = 0,col = "grey")+
  theme_bw()->p6


ggplotly(p7H)

ggplotly(p8H)
```


</div>