---
title: "Mauritius Lipo"
author: "nminaee"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---
# Loading packages 
```{r setup, echo=FALSE, warning = FALSE, message=FALSE}
devtools::load_all("~/nmr-spectra-processing/")
#devtools::install_github("phenological/mva-plots")
library(fusion)
library(dplyr)
library(readr)
library(reshape2)
library(ggplot2)
library(ggrepel)
library(tibble)
library(ggpubr)
library(mva.plots)
library(caret)
library(tidyr)
```


# Loading Mauritius data
```{r echo=FALSE, warning = FALSE, message=FALSE}
setwd("~/Desktop/Diabetes_Obesity/Busselton")# laptop

# reading in Mauritius metadata (temp anno data from RM (06/03/2024))
anno <- read.csv("temp_mauritius_annotation.csv", header = TRUE)

# reading in Mauritius Lipo data element
lipo.daE<-local(get(load("covid19_mauritius_SER_COVr22_LIPO.daE")))

# retrieving the individual lipoprotein data
lipo <- lipo.daE@.Data
colnames(lipo) <- lipo.daE@varName

dim(lipo)
# [1] 272 112

# retrieving the sample annotation
lipo_an<-lipo.daE@obsDescr[[1]] 

# creating sourceID column to match with Survey.No column from anno (metadata)
lipo_an$sourceID<-as.numeric(sapply(strsplit(lipo_an$UUID,"_"),"[",2))

# remove LTR
idx<-which(lipo_an$sampleType %in% c("sample")) # 254 samples - meaning 18 LTR will be taken out
lipo<-lipo[idx,]
lipo_an<-lipo_an[idx,]

#save(lipo, lipo_an, file="MAU_lipo.rda")
```


# Matching metadata with lipo data after LTRs are removed 
```{r echo=FALSE, warning = FALSE, message=FALSE}
idx<-which(anno$Survey.No %in% lipo_an$sourceID) #253
anno<-anno[which(anno$Survey.No%in% lipo_an$sourceID),]
idx<-which(lipo_an$sourceID %in% anno$Survey.No) #253
lipo_an<-lipo_an[idx,]
lipo<-lipo[idx,]
anno<-anno[match(lipo_an$sourceID, anno$Survey.No),] #253

#save(anno, file="MAU_metadata.rda")
```


# Checking NAs in DM column
```{r}
#Survey.No == 17 needs the DM status changed to NDM. 
anno$DM[which(anno$Survey.No=="17")] ="NDM"
```

# Selecting relevant columns from metadata & bind to lipo data 
```{r}
# remember Survey.No == sourceID
annoToBind <- anno %>% select(Survey.No, Med_T2DM, Med_T1DM, DM, Gender, Age, BMI)
lipoMeta <- cbind(lipo, annoToBind)
```

# Creating a new column to identify those with DM + take med and DM + no med. 
```{r}
lipoMeta$Med =""
lipoMeta$Med[which(lipoMeta$DM == "DM" & lipoMeta$Med_T2DM == "No" & lipoMeta$Med_T1DM == "No")] = "DM_NoMed"
lipoMeta$Med[which(lipoMeta$DM == "DM" & lipoMeta$Med_T2DM == "YES" & lipoMeta$Med_T1DM == "No")] = "DM2_Med"
lipoMeta$Med[which(lipoMeta$DM == "DM" & lipoMeta$Med_T2DM == "No" & lipoMeta$Med_T1DM == "YES")] = "DM1_Med"
lipoMeta$Med[which(lipoMeta$DM == "DM" & lipoMeta$Med_T2DM == "YES" & lipoMeta$Med_T1DM == "YES")] = "DM1_2_Med"
lipoMeta$Med[which(lipoMeta$DM == "NDM" & lipoMeta$Med_T2DM == "No" & lipoMeta$Med_T1DM == "No")] = "NDM"
lipoMeta$Med[which(lipoMeta$DM == "NDM" & lipoMeta$Med_T2DM == "YES" & lipoMeta$Med_T1DM == "No")] = "NDM_Med_DM2" # for checking only - just in case
lipoMeta$Med[which(lipoMeta$DM == "NDM" & lipoMeta$Med_T2DM == "No" & lipoMeta$Med_T1DM == "YES")] = "NDM_Med_DM1" # for checking only - just in case
lipoMeta$Med[which(lipoMeta$DM == "NDM" & lipoMeta$Med_T2DM == "YES" & lipoMeta$Med_T1DM == "YES")] = "NDM_Med_DM1_2" # for checking only - just in case


table(lipoMeta$Med)
 # DM_NoMed DM1_2_Med   DM2_Med       NDM 
 #       51        30        65       107
```

NOTE: (confirmed with RM 06/03/2024)
- Mauritius cohort did not specify if they have T1DM or T2DM. 
- DM_NoMed most likely is T2DM. 


# Creating a new column for BMI category
```{r}
lipoMeta$BMI_cut =""
lipoMeta$BMI_cut[which(lipoMeta$BMI<18.49)] = "Underweight"
lipoMeta$BMI_cut[which(lipoMeta$BMI>=18.49 &lipoMeta$BMI<25 )] = "Healthy weight"
lipoMeta$BMI_cut[which(lipoMeta$BMI>=25 &lipoMeta$BMI <30 )] = "Overweight"
lipoMeta$BMI_cut[which(lipoMeta$BMI>=30)] = "Obese"
```

# Creating a new column for comb of BMI category and DM
```{r}
lipoMeta$comb = ""
lipoMeta$comb = gsub("\\s","", paste(lipoMeta$BMI_cut,"_", lipoMeta$DM))
```

```{r}
#save(lipoMeta, file="MAU_lipo_meta_06032024.rda")
#write.csv(lipoMeta, file="MAU_lipo_meta_06032024.csv", row.names = F)
```


```{r}
lipoMeta$Gender[which(lipoMeta$Gender == "Male")] = "M"
lipoMeta$Gender[which(lipoMeta$Gender == "Female")] = "F"
```



```{r}
bus_sau <- read.csv("Bust1_Sau_Healthy_DM2_28112023.csv", header=TRUE)

# renaming and adding columns of bus_sau to be consistent with Mauritius
names(bus_sau )[names(bus_sau ) == "age"] <- "Age"
names(bus_sau )[names(bus_sau ) == "bmi"] <- "BMI"
names(bus_sau )[names(bus_sau ) == "sex"] <- "Sex"
names(bus_sau )[names(bus_sau ) == "BMI.weight"] <- "BMI_cut"

bus_sau$overall_DM =""
bus_sau$overall_DM[which(bus_sau$DM %in% c("DM1", "DM2"))] = "DM"
bus_sau$overall_DM[which(bus_sau$DM == "NO DM")] = "NDM"

bus_sau$overall_comb = ""
bus_sau$overall_comb = gsub("\\s","", paste(bus_sau$BMI_cut,"_", bus_sau$overall_DM))

bus_sau <- bus_sau %>% select(TPTG:sampleID, Age, Sex, BMI, BMI_cut, overall_DM, overall_comb)


# renaming some columns of lipoMeta to be consistent with Busselton and Saudi
names(lipoMeta)[names(lipoMeta) == "Survey.No"] <- "sampleID"
names(lipoMeta)[names(lipoMeta) == "Gender"] <- "Sex"
names(lipoMeta)[names(lipoMeta) == "DM"] <- "overall_DM"
names(lipoMeta)[names(lipoMeta) == "BMI.weight"] <- "BMI_cut"
names(lipoMeta)[names(lipoMeta) == "comb"] <- "overall_comb"

lipoMeta <- lipoMeta %>% select(TPTG:sampleID, Age, Sex, BMI, BMI_cut, overall_DM, overall_comb)

bsm <- rbind(bus_sau, lipoMeta)

bsm$overall_comb[which(bsm$overall_comb == "Healthyweight_NDM")] = "Healthy_NDM"
bsm$overall_comb[which(bsm$overall_comb == "Healthyweight_DM")] = "Healthy_DM"

write.csv(bsm, file = "BUS_SAU_MAU_orig_lipo_meta_06032024.csv", row.names = FALSE)

#add cohort column manually to the csv file and read it back to R. 
orig_bsm <- read.csv("BUS_SAU_MAU_orig_lipo_meta_06032024.csv", header = T) # with an added cohort column

```


# Correcting for age, sex, cohort - ALL Mauritius incl DM_NoMed
```{r}
# running linear model for each of the 112 lipoprotein 
# we want to get the residuals because the residuals are the remaining of lipoprotein's effect after the gender effect is removed. 
# steps: 
# 1. linear model with gender as predictor is fitted 
# 2. extract the residuals 
# 3. use the residuals as the new lipoprotein matrix 

lipoList <- colnames(bsm[, 1:112])
ann <- bsm[, 113:ncol(bsm)]

# converting sex from Male/Female to 0 (Female) and 1 (Male).
bsm$sex01 <- ifelse(bsm$Sex == "F", 0, 1)

# converting cohort from Busselton/Saudi to 0 (Busselton), 1 (Saudi), 2 (Mauritius)
bsm$co012 =""
bsm$co012[which(bsm$cohort == "BUSSELTON")] = 0
bsm$co012[which(bsm$cohort == "SAUDI")] = 1
bsm$co012[which(bsm$cohort == "MAURITIUS")] = 2

# running lm
out <- vector('list', length(lipoList))
resid <- vector('list', length(lipoList))

for (i in seq_along(lipoList)){
  out[[i]] <- lm(paste(lipoList[i],  '~', 'sex01', '+', 'Age', '+', 'co012'),
               data = bsm)
  resid[[i]] <- out[[i]]$residuals
}

# bind all the residuals together to form a matrix 
resid <- do.call("cbind", resid)
colnames(resid) <- colnames(bsm[, 1:112])
class(resid)
#[1] "matrix" "array"
dim(resid)
#[1] 2411 112

# adding grand(overall) median for each lipo back to the residual to recover the scale of the value
grandMedian <- as.list(apply(bsm[, 1:112],2,median))
grandMedian <- apply(bsm[, 1:112],2,median)
length(grandMedian)
#[1] 112
class(grandMedian)
#[1] "numeric"

# making grandMedian shape/dimension the same as resid array so I can add them together to make corrected lipo dataframe/matrix
testrep <- t(matrix(rep(list(grandMedian), times=112), ncol=2411))
testrep2 <- t(do.call(cbind, testrep))
dim(testrep2)
#[1] 2411 112
grandMedianrep <- testrep2

cl <- resid + grandMedianrep # cl = corrected lipo
dim(cl)
#[1] 2411 112

write.csv(cl, file = "BUS_SAU_MAU_correctedLipo_06032024.csv", row.names = F)

bsm <- cbind(cl, ann)
dim(bsm)
#[1] 2411  120

table(bsm$cohort)
# BUSSELTON MAURITIUS     SAUDI 
#      1976       253       182 

bsm$overall_comb[which(bsm$overall_comb == "Healthyweight_NDM")] = "Healthy_NDM"
bsm$overall_comb[which(bsm$overall_comb == "Healthyweight_DM")] = "Healthy_DM"

table(bsm$cohort, bsm$overall_comb)
  #           Healthy_DM Healthy_NDM Obese_DM Obese_NDM Overweight_DM Overweight_NDM
  # BUSSELTON         18         502       81       480            55            834
  # MAURITIUS         32          38       48        21            65             40
  # SAUDI              6          57       44        51             5             19
  #          
  #           Underweight_DM Underweight_NDM
  # BUSSELTON              0               6
  # MAURITIUS              1               8
  # SAUDI                  0               0

write.csv(bsm, file = "BUS_SAU_MAU_correctedLipo_plus_origMeta_06032024.csv", row.names = F)
```
